---
title: "Project_S2"
author: "Xiaohong Deng"
date: "2024-03-02"
output: html_document
---

# Project: Tesla Stock Analysis-Insights from Historical Data Analysis

## Section2: Hypothesis Testing

### 1- Research Question and Relevant Attributes

**Solution:**

Research Question: Is there a statistically significant difference in the average "change_percent" of Tesla's stock before and after March 11, 2020, coinciding with the onset of the COVID-19 pandemic and the Russian invasion of Ukraine?

Relevant Attributes:

-   'change_percent': The daily percentage change in Tesla's stock price.
-   'date': The date of the observation, used to split the dataset into two groups.

### 2- Data Preparation

**Solution:**

Before conducting the hypothesis test, we need to prepare the dataset by adjusting for stock splits and removing rows with missing values in 'change_percent' and 'avg_vol_20d'.

```{r}
#load the updated Tesla dataset
df <- read.csv("tesla_updated.csv")

#convert date column to Date format
df$date <- as.Date(df$date)

#split the dataset into two groups: before and after March 11, 2020
df_before <- df[df$date < as.Date("2020-03-11"), "change_percent"]
df_after <- df[df$date >= as.Date("2020-03-11"), "change_percent"]

#select a random sample to ensure both groups have the same size
sample_size <- min(length(df_before), length(df_after))

df_before_sample <- sample(df_before, size = sample_size)
df_after_sample <- sample(df_after, size = sample_size)
```

### 3- Hypothesis Testing

**Solution:**

H0: There is no significant difference in the average "change_percent" before and after August 31, 2020, considering Tesla's stock splits.

H1: There is a significant difference in the average "change_percent" before and after August 31, 2020, considering Tesla's stock splits.

Test Method: Paired t-test We choose a paired t-test because we are comparing the means of two related groups (before and after the split) with the same subjects. This test is appropriate when the data points in each group are related or matched pairs.

Performing the Paired t-test:

```{r}
#perform the paired t-test with the same size samples
t_test_result <- t.test(df_before_sample, df_after_sample, paired = TRUE)

#check the results
t_test_result
```

Since the p-value is greater than the significance level of 0.05, we fail to reject the null hypothesis(H0). Therefore, we can conclude that there is not a significant difference in the average "change_percent" before and after March 11, 2020, coinciding with the onset of the COVID-19 pandemic and the Russian invasion of Ukraine.

### 4- Rationale Behind Test Selection

**Solution:**

The paired t-test was chosen for the following reasons:

-   Assumption of Normality: The t-test assumes that the differences between pairs are normally distributed, which is often reasonable for large sample sizes.

-   Matched Pairs: We are comparing the same subjects (Tesla's stock performance) before and after a specific event, making it a paired analysis.

-   Same Size Samples: By randomly selecting the same number of observations for both groups, we ensure that the samples have equal size.

These reasons support the selection of the paired t-test, providing a statistical method to assess the significance of the observed difference in Tesla's stock performance before and after March 11, 2020.
