---
title: "Project_S2"
author: "Xiaohong Deng"
date: "2024-03-02"
output: html_document
---

# Project: Tesla Stock Analysis-Insights from Historical Data Analysis

## Section2: Hypothesis Testing

The stock market operates as a complex ecosystem where a diverse range of investors, both large and small, independently make decisions across various investments. It appears as an organized system guided by an invisible force. Market experts contend that specific indicators, such as the Consumer Confidence Index, wars, changes in monetary policy, natural disasters, and corporate performance reports, can help explain the fluctuations in the market(Hall, 2023). So, in hypothesis testing, we'd like to check if the Russian invasion of Ukraine impacts Tesla's stock in either the long-term period or the short-term period.

### 1- Research Question and Relevant Attributes

**Solution:**

Research Question 1(long-term): Is there a statistically significant difference in the average "change_percent" of Tesla's stock before and after Feb 24, 2022, coinciding with the onset of the Russian invasion of Ukraine？

Research Question 2(short-term): Is there a statistically significant difference in the average "change_percent" of Tesla's stock before and after Feb 24, 2022, within a 30-day period coinciding with the onset of the Russian invasion of Ukraine

Relevant Attributes:

-   'change_percent': The daily percentage change in Tesla's stock price.
-   'date': The date of the observation, used to split the dataset into two groups.

### 2.1- Data Preparation for Research Question 1(long-term)

**Solution:**

```{r}
#load the updated Tesla dataset
df <- read.csv("tesla_updated.csv")

#convert date column to Date format
df$date <- as.Date(df$date)

#split the dataset into two groups: before and after Feb 24, 2022
df_before <- df[df$date < as.Date("2022-02-24"), "change_percent"]
df_after <- df[df$date >= as.Date("2022-02-24"), "change_percent"]

#select a random sample to ensure both groups have the same size
sample_size <- min(length(df_before), length(df_after))

df_before_sample <- sample(df_before, size = sample_size)
df_after_sample <- sample(df_after, size = sample_size)
```

### 2.2- Data Preparation for Research Question 2(Short-term)

**Solution:**

```{r}
# Filter the dataset to get the previous 30 days' data before Feb 24, 2022
df_before2 <- df[df$date < as.Date("2022-02-24") & df$date >= as.Date("2022-01-11"), "change_percent"]

# Filter the dataset to get the next 30 days' data after Feb 24, 2022
df_after2 <- df[df$date >= as.Date("2022-02-24") & df$date <= as.Date("2022-04-06"), "change_percent"]

```

```{r}
#length(df_before2)
length(df_before2)
```

```{r}
length(df_after2)
```

### 3.1 - Hypothesis Testing for Research Question 1(long-term)

**Solution:**

H0: There is no significant difference in the average "change_percent" before and after Feb 24, 2024, coinciding with the onset of the Russian invasion of Ukraine.

H1: There is a significant difference in the average "change_percent" before and after Feb 24, 2022, coinciding with the onset of the Russian invasion of Ukraine.

Test Method: Paired t-test We choose a paired t-test because we are comparing the means of two related groups (before and after the split) with the same subjects. This test is appropriate when the data points in each group are related or matched pairs.

Performing the Paired t-test:

```{r}
#perform the paired t-test with the same size samples
t_test_result <- t.test(df_before_sample, df_after_sample, paired = TRUE)

#check the results
t_test_result
```

Since the p-value is greater than the significance level of 0.05, we fail to reject the null hypothesis(H0). Therefore, we can conclude that there is not a significant difference in the average "change_percent" before and after Feb 24, 2022, coinciding with the onset of the Russian invasion of Ukraine.

### 3.2 - Hypothesis Testing for Research Question 2(short-term)

**Solution:**

H0: There is no significant difference in the average "change_percent" before and after Feb 24, 2022 within a 30-day period, coinciding with the onset of the Russian invasion of Ukraine.

H1: There is a significant difference in the average "change_percent" before and after Fbe 24, 2022 within a 30-day period, coinciding with the onset of the Russian invasion of Ukraine.

Test Method: Paired t-test We choose a paired t-test because we are comparing the means of two related groups (before and after the split) with the same subjects. This test is appropriate when the data points in each group are related or matched pairs.

Performing the Paired t-test:

```{r}
#perform the paired t-test with the same size samples
t_test_result2 <- t.test(df_before2, df_after2, paired = TRUE)

#check the results
t_test_result2

```

Since the p-value is less than the significance level of 0.05, we reject the null hypothesis(H0). Therefore, we have sufficient evidence to conclude that there is a significant difference in the average "change_percent" before and after Feb 24, 2022, within a 30-days period, coinciding with the onset of the Russian invasion of Ukraine.

*In Summary:* Based on the hypothesis testing results for Research Question 1 (long-term) and Research Question 2 (short-term), we can summarize the expected impact of the Russian invasion of Ukraine on Tesla's stock:

-   The Russian invasion of Ukraine is not expected to have a significant long-term impact on Tesla's stock, as indicated by the hypothesis testing for the long-time period.
-   However, there is evidence to suggest that the invasion will cause a significant short-term impact on Tesla's stock, as indicated by the hypothesis testing for the short-term period. Within a 30-day period following the invasion, Tesla's stock is expected to experience notable changes in its "change_percent" values.

### 4- Rationale Behind Test Selection

**Solution:**

The paired t-test was chosen for the following reasons:

-   Assumption of Normality: The t-test assumes that the differences between pairs are normally distributed, which is often reasonable for large sample sizes.

-   Matched Pairs: We are comparing the same subjects (Tesla's stock performance) before and after a specific event, making it a paired analysis.

-   Same Size Samples: By randomly selecting the same number of observations for both groups, we ensure that the samples have equal size.

These reasons support the selection of the paired t-test, providing a statistical method to assess the significance of the observed difference in Tesla's stock performance before and after Feb 24, 2022.
